Follow-up Samples

### 2017-01-09 submission/manuscript_outline_20161024.Rmd
update text with merge
update with fig 2 to 6 swtich

***Bacterial Community and Fit Changes before and after Treatment***
Based on thetayc distance metrics, comparing the initial to the follow up samples, there was no difference between the adenoma and carcinoma groups (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`) [Figure 1a]. There was a difference in FIT between initial and follow up samples with the carcinoma group having a significant decrease in FIT versus the adenoma group (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`) [Figure 1b].  Although the thetayc distance metric change was similar between adenoma and carcinoma the directionality of the change was significant in the cancer group between initial and follow up (P-value = `r format(beta_summary['Carcinoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) but not for the adenoma group (P-value = `r format(beta_summary['Adenoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  This change can be visualized using an NMDS [Figure 2]. When all follow up samples were compared to each other there was no significant overall difference between them (P-value = `r format(beta_summary['All', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  There was no significant difference between initial and follow up samples for observed OTUs, Shannon diversity, or evenness after correction for multiple comparisons [Table S1].  Time of follow up sample from initial sampling, did not have a significant difference between adenoma and carcinoma (uncorrected P-value = `r format(time_pvalues["thetayc_change", "uncorrected_p_value"], digits = 3)`).

@@

## Discussion
In our training set we show that the overall community structure as measured by different alpha diversity metrics, shows very little change between controls and those with either adenoma or carcinoma [Table S1].  With respect to our test set there was very little difference in magnitude of change in the thetayc distance metric between those with adenoma or carcinoma [Figure 1a].  In contrast, FIT had a large change in the initial and follow up samples in the carcinoma group versus the adenoma [Figure 1b].  An NMDS showed that there was very little observable change between initial and follow up for the adenoma group but there was one for the carcinoma group [Figure 2].  This cursory information is suggestive that treatment of carcinoma, had the largest response.

@@

**Figure 1: Change in Thetayc and Fit between initial and follow up in adenoma or carcinoma group.**  A) No significant difference was found between the adenoma and carcinoma group for thetayc (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`).  B) A significant difference was found between the adenoma and carcinoma group for FIT (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`).

**Figure 2: NMDS of the Overall Bacterial Community Changes.**  A) NMDS of the intial and follow up samples for the Adenoma group.  B) NMDS of the initial and follow up samples for the Carcinoma group.

@@

**Figure 6: Previously Associated CRC Bacteria in Initial and Follow up Samples.**  A) Carcinoma initial and follow up samples. There was a significant difference in initial and follow up sample for the OTUs classfied as Peptostreptococcus stomatis (P-value = `r format(pvalue_adn_crc_summary["pept", "crc_BH"], digits = 3)`) and Porphyromonas asaccharolytica (P-value = `r format(pvalue_adn_crc_summary["porp", "crc_BH"], digits = 3)`).  B) Adenoma initial and follow up samples.  There were no significant differences between initial and follow up.

>>>>

***Bacterial Community and Fit Changes before and after Treatment***
Based on thetayc distance metrics, comparing the initial to the follow up samples, there was no difference between the adenoma and carcinoma groups (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`) [Figure 1a]. There was a difference in FIT between initial and follow up samples with the carcinoma group having a significant decrease in FIT versus the adenoma group (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`) [Figure 1b].  Although the thetayc distance metric change was similar between adenoma and carcinoma the directionality of the change was significant in the cancer group between initial and follow up (P-value = `r format(beta_summary['Carcinoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) but not for the adenoma group (P-value = `r format(beta_summary['Adenoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  This change can be visualized using an NMDS [Figure 1c & 1d]. When all follow up samples were compared to each other there was no significant overall difference between them (P-value = `r format(beta_summary['All', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  There was no significant difference between initial and follow up samples for observed OTUs, Shannon diversity, or evenness after correction for multiple comparisons [Table S1].  Time of follow up sample from initial sampling, did not have a significant difference between adenoma and carcinoma (uncorrected P-value = `r format(time_pvalues["thetayc_change", "uncorrected_p_value"], digits = 3)`).

@@

## Discussion
In our training set we show that the overall community structure as measured by different alpha diversity metrics, shows very little change between controls and those with either adenoma or carcinoma [Table S1].  With respect to our test set there was very little difference in magnitude of change in the thetayc distance metric between those with adenoma or carcinoma [Figure 1a].  In contrast, FIT had a large change in the initial and follow up samples in the carcinoma group versus the adenoma [Figure 1b].  An NMDS showed that there was very little observable change between initial and follow up for the adenoma group but there was one for the carcinoma group [Figure 1c & 1d].  This cursory information is suggestive that treatment of carcinoma, had the largest response.

@@

**Figure 1: Changes and differences between the adenoma or carcinoma group.**  A) No significant difference was found between the adenoma and carcinoma group for thetayc (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`).  B) A significant difference was found between the adenoma and carcinoma group for FIT (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`). C) NMDS of the intial and follow up samples for the Adenoma group.  D) NMDS of the initial and follow up samples for the Carcinoma group. For C) and D) the teal represents initial samples and the pink represents follow up samples.

**Figure 2: Previously Associated CRC Bacteria in Initial and Follow up Samples.**  A) Carcinoma initial and follow up samples. There was a significant difference in initial and follow up sample for the OTUs classfied as Peptostreptococcus stomatis (P-value = `r format(pvalue_adn_crc_summary["pept", "crc_BH"], digits = 3)`) and Porphyromonas asaccharolytica (P-value = `r format(pvalue_adn_crc_summary["porp", "crc_BH"], digits = 3)`).  B) Adenoma initial and follow up samples.  There were no significant differences between initial and follow up.

@@
